---
title: "dplyr and ggplot2"
author: "Matt Thomas"
date: "January 31, 2018"
output: ioslides_presentation
---

## The tidyverse

"The tidyverse is an opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structures." (https://www.tidyverse.org/)

> * dplyr
> * ggplot2
> * stringr
> * lubridate
> * more...

## dplyr

* pipes
* Good for data manipulation
  * verbs

## dplyr verbs

hflights data set: flights from Houston in 2011
  
```{r}
library(hflights)
hflights
```

## glimpse
With dplyr loaded:
```{r include=FALSE}
library(dplyr)
```
```{r}
glimpse(hflights)
```

## Pipes: %>%

* allow nested functions to be written in the order that you might think about them

## Linear Model
```{r}
summary(lm(DepDelay~DepTime, data=hflights))
```

## Version 2
```{r}
mod1 <- lm(DepDelay~DepTime, data=hflights)
summary(mod1)
```

## With pipes
```{r}
hflights %>% lm(DepDelay~DepTime, data=.) %>% summary()
```

## dplyr verbs

* select()
* filter()
* mutate()
* arrange()
* group_by()
* rename()
* and more - e.g. for merging data frames


## pipes with dplyr verbs

```{r}
hflights %>% 
  filter(DayOfWeek==1) %>% 
  group_by(UniqueCarrier) %>% 
  summarize(MeanDistance=mean(Distance))
```

## ggplot: layered graphics

```{r include=FALSE}
library(ggplot2)
```


```{r}
ggplot(aes(x = DepTime, y = DepDelay), data=hflights)
```

## Adding layers

```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepTime, y = DepDelay), data=hflights)+geom_point()
```

## More layers

```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepTime, y = DepDelay, color=UniqueCarrier), data=hflights)+
  geom_point()
```

## More layers

```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepTime, y = DepDelay), data=hflights)+
  geom_point()+geom_smooth(method="lm", se=F)
```


## Different geoms
```{r}
ggplot(aes(x = DepDelay), data=hflights)
```

## Layers
```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepDelay), data=hflights)+geom_histogram()
```


## Theme layer
```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepDelay), data=hflights)+
  geom_histogram()+
  theme_bw()
```

## Theme layer

```{r include=FALSE}
library(ggthemes)
```


```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepDelay), data=hflights)+
  geom_histogram()+
  theme_economist()
```

## Theme layer

```{r message=FALSE, warning=FALSE}
ggplot(aes(x = DepTime, y = DepDelay, color=UniqueCarrier), data=hflights)+
  geom_point()+
  theme_fivethirtyeight()
```

## Lots of layers

* boxplots
* bargraphs
* error bars
* legends
* rectangles
* curves
* "jitter"

## Upshot

* Packages extend the abilities of R to do *many* things
* dplyr and ggplot are two packages which extend the capabilities of R in a particularly structured way

